---
title: "Archive Sheet – Preparation of RPG Data"
author: "Loïc Henry"
date: "2026-05-06"
output: html_document
---

# Data Overview

This work draws on the **Registre Parcellaire Graphique (RPG)** geographic database, produced by the *Agence de Services et de Paiement* (ASP) since 2007 and publicly available at <https://geoservices.ign.fr/rpg>.

These annual datasets contain the graphic boundaries of agricultural parcels along with their primary crop type, covering the entire French territory including Mayotte and Saint-Martin, but excluding Saint-Barthélemy.

In addition to the RPG parcel data, the workflow relies on reference tables. One essential reference table maps all crop codes used in the dataset.

# Replication Instructions: Preparing the Registre Parcellaire Graphique (RPG) Data

The main objetive of the data preparation is to aggregate the plot level data at a larger spatial scale, either commmune-level or regional scale. The data preparation pipeline transforms raw RPG files into analysis-ready datasets through the following steps.

## Step 1 — Download Raw Data

**Script:** `00_DownloadData.R`

Downloads the raw RPG files for the analysis period (2007–2023) from the IGN public data service.
One individual download consist of one year-region vintage. This step thus downloads multiple
files which are then stored in `data/raw/` for subsequent processing.

Note that this script is no longer valid as IGN has changed its data delivery process.

Once you've downloaded all data folders by region and year, you can then move to the next
step.

## Step 2 — Aggregate Parcel Data

### Step 2a — Regional aggregation

**Script:** `01_AggregateData.R`

All available downloaded `shapefile` inputs are aggregated by region and year at the municipal (*commune*) level of detail, or at the regional (*petites régions agricoles*) level. This intermediate aggregation is designed to reduce file sizes and facilitate downstream processing.

Aggregation is done both at the broader crop group level (20-ish crop categories)
and at the finer specific crop type level (about 200 crop types).

**Output:** one `.rds` file per region × year combination.

### Step 2b — National consolidation

**Script:** `02_AssembleData.R`

The `.rds` files produced in Step 2a are merged into a single national dataset for all years. This step includes:

- Cleaning inconsistencies across regional files
- Re-aggregating by crop group
- Producing a homogeneous national panel ready for large-scale analysis

**Output:** `RPG_Aggregated_ALL_wide.rds` (stored in `data/final/`)

# Mapping and Visualizing data (Exploratory)

The following scripts were developed for exploratory work at a local scale (only for one region) and are not part of the main pipeline, but mainly there to run some premiminary exploratory analysis on RPG data.

**Script:** `03_MapCommunes.R` — Annual crop maps at the municipal level. Provides a static snapshot of the agricultural situation in the study region at a given year.

**Script:** `03_MapParcels.R` — Inter-annual comparison of agricultural parcels. Enables spatial analysis of crop changes and land-use dynamics within a given region, producing both static and interactive maps.

# Descriptive Statistics

After running the scripts from `00_XXX`to `02_XXX` above, the final dataset aggregates RPG parcel data at the municipal (or regional) and national levels for all years.

The code block below computes basic summary statistics from the final wide-format dataset.

```{r prepare_environment, include = FALSE}
# Clean memory 
rm(list = ls())
gc()

# Load packages
if (!require("pacman")) install.packages("pacman")
pacman::p_load(dplyr, here, tidyr)

# Define directory structure
dir <- list()
dir$root     <- here()
dir$data     <- here(dir$root, "data")
dir$raw      <- here(dir$data, "raw")
dir$derived  <- here(dir$data, "derived")
dir$final    <- here(dir$data, "final")
dir$script   <- here(dir$root, "script")
dir$output   <- here(dir$root, "output")

# Create directories if they do not already exist
lapply(dir, function(i) dir.create(i, recursive = TRUE, showWarnings = FALSE))
```

```{r}
RPG_wide <- readRDS(here(dir$final, "RPG_Aggregated_ALL_wide.rds"))

cultures_stat <- RPG_wide %>%
  group_by(year) %>%
  summarise(
    N_communes          = n_distinct(insee),
    surf_agri_year      = sum(surf_agri_geo_unit_m2, na.rm = TRUE)
  ) %>%
  mutate(
    mean_surf_agri_communes = surf_agri_year / N_communes,
    mean_surf_agri_tot      = mean(surf_agri_year)
  )

print(cultures_stat)
```